---
title: "AEM: Problem Set 1"
author: "Maxwell Austensen"
date: "September 24, 2016"
output: pdf_document
---

```{r chuck-opts, include=FALSE}
# output_type <- "html" # for Starazer output
output_type <- "latex" # for Starazer output
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


```{r setup}
# Utility functions -------------------------------------------------------

`%S%` <- function(x, y) {
  paste0(x, y)
}

`%notin%` <- Negate(`%in%`)

################################################################################

# Install packages if needed
package_list <- c("stargazer", "knitr", "haven", "labelled", "ICC", "scales", "tidyverse")
new_packages <- package_list[package_list %notin% installed.packages()[,"Package"]]
if(length(new_packages)) install.packages(new_packages)

library(stargazer)
library(knitr)
library(haven)
library(labelled)
library(ICC)
library(scales)
library(tidyverse)

# Set directories
repo_ <- "H:/GitHub/aem/"
ps1_ <- "C:/Users/austensen/Box Sync/aem/ps1/"

################################################################################
```




```{r load-data}
data_raw <- 
  read_stata(ps1_ %S% "Thornton HIV Testing Data.dta") %>% 
  remove_val_labels

names(data_raw) <- names(data_raw) %>% tolower

main_sample <-
  data_raw %>% 
  filter(
    hiv2004 %notin% c(NA, -1),
    !is.na(any),
    !is.na(zone),
    !is.na(age)
  )
```

## Part I: Summary Statistics

```{r}
get_summary <- function(data){
  data %>% 
    summarise(
      `Average Age` = mean(age, na.rm=T),
      `Percentage of Males` = mean(male, na.rm=T)*100,
      `Average Years of Education` = mean(educ2004, na.rm=T),
      `Percentage with HIV` = mean(hiv2004, na.rm=T)*100
    ) %>% 
    kable(digits = 1)
}
```

### 1. 
```{r}
main_sample %>% get_summary
```

***

### 2.
```{r}
main_sample %>% group_by(any) %>% get_summary
main_sample %>% group_by(under) %>% get_summary
```

There are no major differences in the variables betweeen treatment and control groups based on either cash recipt or distance. However, those that received some cash were on average about a year older and those that were under 1.5K from the center had an HIV rate three percentage points high than those further away.

***

### 3.
```{r results = "asis"}
grps <- c("any", "under")
vars <- c("age", "male", "hiv2004")

ttest_results <- 
  data.frame(
    var = character(), 
    P.value = double(), 
    stringsAsFactors=FALSE
  )

for(grp in grps) {
  i <- 1
  for(var in vars){
    result <- t.test(main_sample[[var]] ~ main_sample[[grp]], var.equal = TRUE)$p.value
    
    ttest_results[i,1] <- var
    ttest_results[i,2] <- result
    i <- i+1
  }
  writeLines("#### t-test: group = " %S% grp)
  print(kable(ttest_results, digits = 3))
  writeLines("\n\n\n")
}
```


***

## Part II: Analysis using graphs

### 4.
```{r}
main_sample %>% 
  group_by(any) %>% 
  summarise(
    got_mean = mean(got, na.rm=T),
    got_se = sd(got)/sqrt(n())
  ) %>% 
  ggplot(aes(any, got_mean)) +
  geom_bar(stat = "identity", fill = "#E69F00") +
  geom_errorbar(aes(ymin = got_mean - got_se, ymax = got_mean + got_se), width = 0, size = 1) +
  geom_text(aes(y = got_mean+0.05, label = round(got_mean*100, 1) %S% "%")) +
  scale_y_continuous(labels = percent_format()) +
  scale_x_continuous(breaks = c(0,1), labels = c("No Incentive", "Some Incentive")) +
  ylab("Percentage learning HIV results") + xlab("") +
  ggtitle("Effects of receiving some incentive")
```


***

### 5.
```{r}
main_sample %>% 
  group_by(ti) %>% 
  summarise(
    got_mean = mean(got, na.rm=T),
    got_se = sd(got)/sqrt(n())
  ) %>% 
  ggplot(aes(ti, got_mean)) +
  geom_bar(stat = "identity", fill = "#E69F00") +
  geom_errorbar(aes(ymin = got_mean - got_se, ymax = got_mean + got_se), width = 0, size = 1) +
  geom_text(aes(y = got_mean+0.05, label = round(got_mean*100, 1) %S% "%")) +
  scale_y_continuous(labels = percent_format()) +
  scale_x_continuous(breaks = c(0,50, 100, 200, 300), labels = c("$0", "$0.50", "$1.00", "$2.00", "$3.00")) +
  ylab("Percentage learning HIV results") + xlab("") +
  ggtitle("Effects of the total amount of the incentive")
```

***
\newpage

## Part III: Analysis using linear regression

### 6.
```{r results = "asis"}

any_1 <- lm(got ~ any, data = main_sample)
any_2 <- lm(got ~ any + age + male + educ2004 + mar, data = main_sample)

stargazer(any_1, any_2, type = output_type)
```

\newpage
### 7.
```{r}
group_diff_any <- 
  main_sample %>% 
  group_by(any) %>% 
  summarise(got = mean(got*100, na.rm=T)) %>% 
  spread(any, got) %>% 
  transmute(diff = `1` - `0`) %>% 
  .[[1]]
```

Using a group means comparison, the estimated treatment effect is `r round(group_diff_any, 2)`.
This answer does not differ significantly from the OLS coefficent estimate on treatment.
Since the addition of control variables does not significatly alter the treatment effect estimate, this suggests that the randomization of treatment was successful in balancing the two groups with respect to these other variables.

***

\newpage
### 8.
```{r results = "asis"}
ti_1 <- lm(got ~ ti, data = main_sample)
ti_2 <- lm(got ~ ti + age + male + educ2004 + mar, data = main_sample)

stargazer(ti_1, ti_2, type = output_type)
```


***

### 9.


***

\newpage
## Part IV: Conditional (Heterogeneous) Treatment Effects

### 10.
```{r results = "asis"}
any_male <- lm(got ~ any + male + any*male, data = main_sample)

stargazer(any_male, type = output_type)
```
The estimate for the treatment-male interaction term is `r round(any_male$coefficients[["male"]], 3)`.
It is not statistically significant.
This suggests that there is not differential effect of recieving any cash ammount for men and women.

***

\newpage
### 11.
```{r results = "asis"}
any_educ <- lm(got ~ any + educ2004 + any*educ2004, data = main_sample)

stargazer(any_educ, type = output_type)
```


***

## Part V: Policy Implications

### 12.


***

### 13.


***

\newpage

## Part VI: A Random Sub-Sample

### 14.
```{r}
sample_1000 <- main_sample %>% sample_n(1000)
```

### 15.
```{r results = "asis"}

any_1 <- lm(got ~ any, data = sample_1000)
any_2 <- lm(got ~ any + age + male + educ2004 + mar, data = sample_1000)

stargazer(any_1, any_2, type = output_type)
```


## Part VII: Choosing Sample Size

### 16. 


***

### 17.
```{r}
main_sample %>% 
  filter(!is.na(usecondom04)) %>% 
  ICCbare(site, usecondom04, data = .)
```

***

## Part VIII: Fisher Randomization Test (bonus)

### 18.
```{r}
# data %>% 
#   filter(!is.na(any)) %>% 
#   group_by(any) %>% 
#   summarise(got = mean(got, na.rm=T))
# 
# oneway_test(got ~ factor(any), data, alternative = "greater")
# independence_test(got ~ factor(any), data, 
#                   alternative = "greater", 
#                   distribution = approximate(B = 1000))
# 
# fisher.test(data$got, factor(data$any), alternative = "greater", simulate.p.value = TRUE)

```

***




